---
title: "Netlogo to R Pipeline: Consumer-Resource"
author: "Shannon Carter"
date: "March 11, 2019"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: united
    highlight: tango 
---

# Goal

The objective here is to make a clean reproducible pipeline to transform Netlogo output into a tidy rectangular dataframe. Here, I import behavior space data (i.e., experiment results) from Netlogo to R. The data I've imported here does several 'runs' (multiple treatments and replicates) of an simulation testing effects of phenology on consumer-resource interaction. There is summary data for each run and there's also time series info on size for each turtle/agent. Here, I process the data and make some diagnostic plots to finetune the model.

```{r setup, include = F}
knitr::opts_chunk$set(echo = T, warning = F, message = F)

## Set working directory and clear memory
setwd("C:\\Users\\Shannon\\Desktop\\GitHub Repos\\phenology-abm")
rm(list = ls(all = T))

## Load required packages
library(tidyverse)
library(reshape2)
library(RColorBrewer)
library(lme4)
library(wesanderson)
#library(RNetLogo)    # maybe worthwhile looking into this, so far can't figure it out

## Load universal plotting elements
mytheme <- theme(panel.background = element_blank(),
                 panel.grid.minor = element_blank(), 
                 panel.grid.major = element_blank(),
                 axis.text  = element_text(size = rel(1.3), colour = "black"),
                 axis.title = element_text(size = rel(2.0)),
                 axis.line  = element_line(colour = "black"))

```

# Import Data

First, be sure you export behavior space data from Netlogo in "table" format. When the raw data is exported as a spreadsheet, it comes in a really gnarly format. R can't read it because it's not even close to rectangular, big header, mix of space delimited and bracket delimited. It's a mess. The table is still a bit messy, but much easier to wrangle.

```{r load table data}
## first test with 1 consumer and resource. Size scaled consumption
#testa <- read.table("bs_CR1.csv",
#                   header = T,  
#                   sep = ',',        # define the separator between columns
#                   skip = 6,         # 6 header rows we don't need
#                   quote = "\"",     # correct the column separator
#                   fill = T)         # add blank fields if rows have unequal length

## second test with 1 consumer and resource. size-independent consumption
test <- read.table("bs_CR4.csv",
                   header = T,  
                   sep = ',',        # define the separator between columns
                   skip = 6,         # 6 header rows we don't need
                   quote = "\"",     # correct the column separator
                   fill = T)         # add blank fields if rows have unequal length
# [,1:12] are parameter values (some consistent across all, some representing trts)
# [,13:ncol] are the output we designated
str(test)  

```

# Inspect Data

First, let's rename columns and drop some that we don't need. I keep some columns that are the same across all model runs because it will be useful to call them later, especially since different runs of the experiment might have differetn values. Some columns will always be the same, or are irrelevant. These I drop. 

For this model, I had two turtle types (fishes and dflies), and varied the timing of hatching and densities of each. Outputs I collected from Netlogo were number of survivors of each turtle type and a vector of sizes for each individual turtle.

```{r}
# select and rename columns. new_name = old name. 
# renaming everything I keep, so can rename with select
test <- test %>% 
  select(run_num = X.run.number.,                  # each run number is an experimental unit
         total_time = X.step.,                     # length of time the experiment ran
         n_fish = n.fishes,                        # two types of "turtles" here: fish & dfly
         sync_fish = var.hatch.fishes,             # amount of variation in ind hatching timing
         surv_fish = n.meta.fishes,                # number that 'metamorphed', i.e., size >10
         sizes_fish = X.fish.size.list..of.fishes) # list of size for each ind for each time step

# Extract parameter values-- these are the 'treatments' and inputs of the model for each run
# I'll rejoin the output data with params later, after some summarizing/processing
# Here, add qualitative levels for the treatment parameters. This will be different for each BS
params <- test[,1:5]

```

# Parse Individuals

The 'sizes_fish' columns have a time series of size for each individual (so each 'sizes_fish' column has 50 individuals x 150 time steps values). We ultimately want one row per individual per time step. Let's first separate by individual. 

```{r}

# First, separate so that each individual has it's own column
test <- test %>% 
  separate(sizes_fish,                                      # separate sizes_fish
           into = paste("f_", c(1:max(test$n_fish)), sep = ""),
           #into = as.character(c(1:max(test$n_fish))),  # levels for new var
           sep = "]")                # every ] marks a new ind                                       
# some cells are left blank and not NA (because of "]]" ending), so fill those with NA
test[test==""] <- NA

# Next, I'll gather individuals so that they appear in rows

# Select only unique fish individuals and gather to long format
test_fish <- test %>%
  select(run_num, f_1:f_100) %>% 
  unique(.) %>% 
  gather(f_1:f_100, key = "fish_id", value = 'size_fish') %>% 
  arrange(run_num)

# remove NA cases, i.e. for runs where n_fish < max(n_fish)
test_fish <- test_fish[complete.cases(test_fish), ]

# Remove the brackets
test_fish$size_fish <- gsub("[", "", test_fish$size_fish, fixed = T)
```

# Parse Time Series

Now, do the same separate and gather operations to parse the time vector and put in long format.

```{r}

# Separate the size time series- each space represents a new time point
test_fish <- test_fish %>% 
  separate(size_fish, into = as.character(c(0:149)), sep = " ") %>% 
  select(-c(as.character(0))) # I drop this because there's only a value for ind 1

# Gather to long format, so we now have one line per fish per time step
test_fish <- test_fish %>% 
  group_by(run_num) %>% 
  gather(as.character(1:149), key = "time", value = 'current_size_fish') %>%
  arrange(run_num, fish_id)

```

# Compile & Tidy

Now, rejoin with the treatment information, and write a csv. This csv has one line per individual per time step. So nrow should be max(run_num) x (ndfly + nfish) x total_time
```{r}

# Rejoin with treatment identifiers for each run number
test <- left_join(test_fish, params, by = "run_num")
ind_time <- test %>%
  select(run_num, n_fish, sync_fish,
         surv_fish, time, fish_id, current_size_fish) %>% 
  rename(size = current_size_fish, id = fish_id)

write.csv(ind_time, "individual_time_CR.csv")

```


## Summarize by Individual

Now, I want to make some summarised datasets. First with one row per individual, then with one row per run number 

```{r}
ind_time <- read.csv("individual_time_CR.csv", header = T)

# producing a lot of NAs
ind <- ind_time %>% 
  group_by(run_num, id, sync_fish, surv_fish, n_fish) %>%
  summarise(max_size = min(max(size), 11),
            meta = if_else(max_size == 11, 1, 0),
            hatch_date = min(time[size > 0]),
            # use base R ifelse here, so we can have 2 data types for T/F
            end_date = ifelse(meta == 1, min(time[size == 12]), min(time[size == max_size])),
            growth_time = end_date - hatch_date)

# On some data runs, number of metamorphs doesn't match raw data. always double check.
check <- ind %>% 
  group_by(run_num, surv_fish, n_fish) %>%
  summarise(no_meta = sum(meta, na.rm = T))
ggplot(check, aes(x = surv_fish, y = no_meta, color = as.factor(n_fish))) +
  geom_point() + theme_bw() +
  geom_abline(slope = 1, intercept = 0) #+ xlim(10, 60) + ylim(10, 60)

```

# Plots

Here are some diagnostic plots to get a quick look at the data. These will give an idea of what parameter space is reasonable and interesting to explore in future behavior space runs.


```{r, echo = F}
ggplot(ind, aes(x = hatch_date)) + theme_bw() +
  geom_density(alpha = 0.5, size = 1.2) +
  facet_grid(. ~ sync_fish) +
  labs(title = "Hatching timing treatments",
       x = "hatching date",
       y = "density of individuals")
```

```{r}
ggplot(ind_time, aes(x = time, y = size, color = id)) +
  #geom_point() +
  geom_smooth(se = F) + theme_bw() +
  facet_wrap(~run_num) +
  theme(legend.position = "none") 
  
ggplot(ind, aes(x = hatch_date, y = growth_time, color = as.factor(meta))) +
  geom_jitter()

# tons of NA-- look into this
ggplot(ind, aes(x = max_size, y = growth_time, color = as.factor(meta))) +
  geom_jitter()
```


```{r, echo = F}
ggplot(ind, aes(x = hatch_date, y = max_size)) +
  geom_point(size = 1, alpha = 0.2, position = 'jitter') +
  stat_smooth() +
  facet_grid(. ~ sync_fish, scales = 'free') + theme_bw() +
  labs(title = "Final size by hatching date",
       x = "hatching date",
       y = "final size")

```

```{r, echo = F}

ggplot(ind, aes(x = hatch_date, y = growth_time)) + theme_bw() +
  geom_point(alpha = 0.65, size = 2) +
  stat_smooth(method = "lm", formula = y ~ poly(x, 2), se = T, size = 2) +
  facet_grid(. ~ sync_fish, scales = "free") + 
  theme(#legend.position = "none",
        strip.text.x = element_text(size = 14),
        strip.text.y = element_text(size = 14)) +
  labs(title = "Growth rate by hatching date",
       x = "hatching date",
       y = "days to metamorphosis")

```

```{r, echo = F}
# Looks like in several cases, there are some individuals hatching on the last day, way later than the others. Check.
ggplot(ind, aes(x = hatch_date, y = meta)) + theme_bw() +
  geom_point(alpha = 0.25) + 
  stat_smooth(method = 'glm', method.args = 'binomial', se = F) +
  facet_grid(. ~ sync_fish, scales = "free") + 
  xlab("hatch date") + ylab("probability of survival") +
  theme(strip.text.x = element_text(size = 14),
        strip.text.y = element_text(size = 14)) +
  labs(title = "Survival probability by hatching date",
       x = "hatching date",
       y = "probability of survival")
```

```{r, echo = F}

ggplot(ind, aes(x = hatch_date, y = meta, color = as.factor(sync_fish))) +
  geom_point(alpha = 0.25) + theme_bw() +
  stat_smooth(method = 'glm', method.args = 'binomial', size = 2, alpha = 0.5, se = F) +
  labs(title = "Probability of survival by hatching date",
       x = "hatching date",
       y = "probability of survival")
```

```{r, echo = F}

## Subset to include only individuals that metamorphed
ind_surv <- subset(ind, subset = (meta == 1))

## Original hatching and metamorph
ggplot(ind_surv, aes(hatch_date)) + theme_bw() +
  # gray: original/imposed hatching synchrony
  geom_density(size = 1, alpha = 0.25, fill = "black", color = "gray39") + 
  # blue: metamorphosis dates of survivors, scaled to day 0
  geom_density(data = ind_surv, size = 1, alpha = 0.5, color = "steelblue4", fill = "steelblue4",
               aes(x = meta_date)) +
  facet_grid(. ~ sync_fish, scales = "free_y") + 
  theme(strip.text.x = element_text(size = 14),
        strip.text.y = element_text(size = 14)) +
  labs(title = "Phenology across ontogeny",
       x = "date",
       y = "density of individuals")

```

