---
title: "Netlogo to R Pipeline"
author: "Shannon Carter"
date: "March 11, 2019"
output: 
  html_document:
  theme: united
---

The goal here is to make a clean reproducible pipeline to transform Netlogo output into a tidy rectangular dataframe. Here, I import behavior space data (i.e., experiment results) from Netlogo to R. Netlogo's behavior space data comes in a wonky format, and the goal is to do all tidying in R (vs. on spreadsheet) so it's easy to feed through new data. The data I've imported here does several 'runs' (multiple treatments and replicates) of an simulation testing effects of phenology on species interactions. There is summary data for each run and there's also time series info on size for each turtle/agent.

```{r setup, include = F}
knitr::opts_chunk$set(echo = T, warning = F)

## Set working directory and clear memory
setwd("C:\\Users\\Shannon\\Desktop\\GitHub Repos\\phenology-abm")
rm(list = ls(all = T))

## Load required packages
library(tidyverse)
library(reshape2)
library(RColorBrewer)
library(lme4)
library(wesanderson)
#library(RNetLogo)    # maybe worthwhile looking into this, so far can't figure it out

## Load universal plotting elements
mytheme <- theme(panel.background = element_blank(),
                 panel.grid.minor = element_blank(), 
                 panel.grid.major = element_blank(),
                 axis.text  = element_text(size = rel(1.3), colour = "black"),
                 axis.title = element_text(size = rel(2.0)),
                 axis.line  = element_line(colour = "black"))

```

## Import Data

First, be sure you export behavior space data from Netlogo in "table" format. When the raw data is exported as a spreadsheet, it comes in a really gnarly format. R can't read it because it's not even close to rectangular, big header, mix of space delimited and bracket delimited. It's a mess. The table is still a bit messy, but much easier to wrangle.

```{r load table data}

test <- read.table("bs2.csv",
                   header = T,  
                   sep = ',',        # define the separator between columns
                   skip = 6,         # 6 header rows we don't need
                   quote = "\"",     # correct the column separator
                   fill = T)         # add blank fields if rows have unequal length

# [,1:12] are parameter values (some consistent across all, some representing trts)
# [,13:ncol] are the output we designated
str(test)  

```

# Select & Rename Variables

First, let's rename columns and drop some that we don't need. I keep some columns that are the same across all model runs because it will be useful to call them later, especially since different runs of the experiment might have differetn values. Some columns will always be the same, or are irrelevant. These I drop. 
```{r}
# select and rename columns. new_name = old name. 
# renaming everything I keep, so can rename with select
test <- test %>% 
  select(run_num = X.run.number.,
         total_time = X.step.,
         n_fish = n.fishes,
         n_dfly = n.dflies,
         sync_dfly = var.hatch.dflies,   
         mean_dfly = mean.hatch.dflies, 
         surv_dfly = n.meta.dflies,
         sync_fish = var.hatch.fishes, 
         surv_fish = n.meta.fishes, 
         sizes_dfly = X.dfly.size.list..of.dflies, 
         sizes_fish = X.fish.size.list..of.fishes)

# extract parameter values-- will need it for rejoining data with params later
# here, it'd also be a good idea to give qualitative/ranked levels to treatments (i.e., high, med, low)
# this will be different for each behavior space
params <- test[,1:9]
params$mean_dfly_num <- params$mean_dfly
params$sync_dfly_num <- params$sync_dfly
params$sync_dfly <- as.factor(params$sync_dfly)
params$mean_dfly <- as.factor(params$mean_dfly)
levels(params$sync_dfly)
levels(params$sync_dfly) <- c("high", "med", "low")
levels(params$mean_dfly) <- c("dfly v early", "dfly early", "same", "fish early", "fish v early")

```

# Parse Individuals

The 'sizes_x' columns have a time series of size for each individual (so each sizes_ column has 50 individuals x 150 time steps values). We ultimately want one row per individual per time step. Let's first separate by individual. 

```{r}

# first, separate so that each individual has it's own column
test <- test %>% 
  separate(sizes_dfly,                                      # separate sizes_dfly
           into = paste("d_", c(1:test$n_dfly), sep = ""),  # levels for new var
           sep = "]") %>%                                   # every ] marks a new ind
  separate(sizes_fish, 
           into = paste("f_", c(1:test$n_fish), sep = ""), 
           sep = "]")

# next, I'll gather individuals so that they appear in rows
# I separate this step by species because it's easier for me to catch errors

## FIRST, DRAGONFLIES 
# select only unique dfly individuals and gather to long format
test_dfly <- test %>%
  select(run_num, d_1:d_50) %>% 
  unique(.) %>% 
  gather(d_1:d_50, key = "dfly_id", value = 'size_dfly') %>% 
  arrange(run_num)

# remove the brackets
test_dfly$size_dfly <- gsub("[", "", test_dfly$size_dfly, fixed = T)

## NEXT, FISH
# select only unique fish individuals and gather to long format
test_fish <- test %>%
  select(run_num, f_1:f_50) %>% 
  unique(.) %>% 
  gather(f_1:f_50, key = "fish_id", value = 'size_fish') %>% 
  arrange(run_num)

# remove the brackets
test_fish$size_fish <- gsub("[", "", test_fish$size_fish, fixed = T)
```

# Parse Time Series

Now, do the same separate and gather operations to parse the time vector and put in long format.

```{r}

## FIRST, DRAGONFLIES 
# separate the size time series- each space represents a new time point
test_dfly <- test_dfly %>% 
  separate(size_dfly, into = as.character(c(0:149)), sep = " ") %>% 
  select(-c(as.character(0))) # I drop this because there's only a value for ind 1

# gather to long format, so we now have one line per dfly per time step
test_dfly <- test_dfly %>% 
  group_by(run_num) %>% 
  gather(as.character(1:149), key = "time", value = 'current_size_dfly') %>%
  arrange(run_num, dfly_id)

## NEXT, FISH
# separate the size time series- each space represents a new time point
test_fish <- test_fish %>% 
  separate(size_fish, into = as.character(c(0:149)), sep = " ") %>% 
  select(-c(as.character(0))) # I drop this because there's only a value for ind 1

# gather to long format, so we now have one line per dfly per time step
test_fish <- test_fish %>% 
  group_by(run_num) %>% 
  gather(as.character(1:149), key = "time", value = 'current_size_fish') %>%
  arrange(run_num, fish_id)

```

# Compile & Tidy

Now, join the 2 species' dataframes together, rejoin with the treatment information, and write a csv. This csv has 1 line per individual per time step. So nrow should be max(run_num) x (ndfly + nfish) x total_time
```{r}

# paste dfs together and remove redundant variables
test <- cbind(test_dfly, test_fish)
test <- test %>% 
  select(-c(run_num1, time1)) 

# this is clunkier than it needs to be, but allows some double checking
test$dfly <- paste(test$dfly_id, test$current_size_dfly)
test$fish <- paste(test$fish_id, test$current_size_fish)
test1 <- test %>% 
  select(run_num, time, dfly, fish)
test2 <- test1 %>% 
  gather(key = "species", "size", dfly, fish)
test2$time <- as.numeric(test2$time)

test2 <- test2 %>% 
  separate(size, into = c("id", "size"), sep = " ") %>% 
  separate(id, into = c("sp", "id"), sep = "_") %>% 
  select(-sp) %>% 
  arrange(run_num, id, time)
test2$size <- as.numeric(test2$size)

# rejoin with treatment identifiers for each run number
test <- left_join(test2, params, by = "run_num")
ind_time <- test %>%
  select(run_num, n_fish, n_dfly, sync_fish, sync_dfly, mean_dfly,
         surv_fish, surv_dfly, time, species,
         id, size)

write.csv(ind_time, "individual_time.csv")

```


## Summarize by Individual

Now, I want to make some summarised datasets. First with one row per individual, then with one row per run number 

```{r}
ind_time <- read.csv("individual_time.csv", header = T)

ind <- ind_time %>% 
  group_by(run_num, species, id, sync_fish, sync_dfly, mean_dfly, surv_dfly, surv_fish) %>%
  summarise(max_size = max(size),
            meta = if_else(max_size >= 10, 1, 0),
            hatch_date = min(time[size > 0]),
            # use base R ifelse here, so we can have 2 data types for T/F
            meta_date = ifelse(meta == 1, min(time[size >= 10]), NA), 
            death_date = ifelse(meta == 0, min(time[size = max(max_size)]), NA),
            growth_time = meta_date - hatch_date)

# currently, number of metamorphs doesn't match raw data. think it might be a >= mismatch. check.
check <- ind %>% 
  group_by(run_num, species, surv_dfly, surv_fish) %>%
  summarise(no_meta = sum(meta))
ggplot(subset(check, subset = (species == "fish")), aes(x = surv_fish, y = no_meta)) +
  geom_point() + theme_bw() +
  geom_abline(slope = 1, intercept = 0)

```

## Summarize by Treatment

Now, make a df with 1 row per unique treatment with means/se's 
```{r}
trt <- params %>% 
  group_by(sync_dfly, mean_dfly, mean_dfly_num, sync_dfly_num) %>% 
  summarise(mean_surv_dfly = mean(surv_dfly),
            mean_surv_fish = mean(surv_fish),
            se_surv_dfly   = sd(surv_dfly)/sqrt(length(surv_dfly)),
            se_surv_fish   = sd(surv_fish)/sqrt(length(surv_fish)))

ind$sync_dfly <- factor(ind$sync_dfly, levels = c("low", "med", "high"))
ind$mean_dfly <- factor(ind$mean_dfly, levels = c("dfly v early", "dfly early", "same", "fish early", "fish v early"))
trt$sync_dfly <- factor(trt$sync_dfly, levels = c("low", "med", "high"))
trt$mean_dfly <- factor(trt$mean_dfly, levels = c("dfly v early", "dfly early", "same", "fish early", "fish v early"))

```

## Plots

# First, a visualization of the treatments
```{r, echo = F}
ggplot(ind, aes(x = hatch_date, color = species, fill = species)) + theme_bw() +
  geom_density(alpha = 0.5, size = 1.2) +
  facet_grid(sync_dfly ~ mean_dfly) +
  scale_color_manual(values = c("#F28335", "#489C92")) +
  scale_fill_manual(values = c("#F28335", "#489C92"))
```

# Final size by hatch date

```{r, echo = F}
plot3 <- ggplot(ind, aes(x = hatch_date, y = max_size, color = species)) +
  geom_point(size = 1, alpha = 0.2, position = 'jitter') +
  stat_smooth() +
  facet_grid(sync_dfly ~ mean_dfly, scales = 'free') + theme_bw() +
  scale_color_manual(values = c("#F28335", "#489C92")) 
plot3

```

# Survival probability based on hatching date

Looks like in 'fish early' scenarios, there are some fish hatching on the last day. Check.

```{r, echo = F}
plot4 <- ggplot(ind, aes(x = hatch_date, y = meta, color = species)) + theme_bw() +
  geom_point(alpha = 0.25) + 
  stat_smooth(method = 'glm', method.args = 'binomial', se = F) +
  facet_grid(sync_dfly ~ mean_dfly, scales = "free") + 
  xlab("hatch date") + ylab("probability of survival") +
  theme(strip.text.x = element_text(size = 14),
        strip.text.y = element_text(size = 14)) +
  scale_color_manual(values = c("#F28335", "#489C92")) 
plot4
```

# Dragonfly survivors by treatment

```{r, echo = F}
plot5 <- ggplot(trt, aes(x = mean_dfly_num, y = mean_surv_dfly, color = sync_dfly)) +
  geom_point(size = 2) + 
  #geom_errorbar(width = 0, aes(max = mean_surv_dfly + se_surv_dfly, min = mean_surv_dfly - se_surv_dfly)) +
  stat_smooth(size = 2, method = "lm", se = T, 
              aes(x = mean_dfly_num, fill = sync_dfly)) +
  xlab("mean arrival") + ylab("number survivors")  +
  scale_fill_manual(values = wes_palette(n = 3, name = "GrandBudapest1")) +
  scale_color_manual(values = wes_palette(n = 3, name = "GrandBudapest1")) +
  theme_bw()
plot5

```

# Phenological synchrony from hatching to metamorphosis

```{r, echo = F}

## Subset to include only individuals that metamorphed
ind_surv <- subset(ind, subset = (meta == 1 & species == "dfly"))

## Original hatching and metamorph
plot6 <- ggplot(ind_surv, aes(hatch_date)) + theme_bw() +
  # gray: original/imposed hatching synchrony
  geom_density(size = 1, alpha = 0.25, fill = "black", color = "gray39") + 
  # blue: metamorphosis dates of survivors, scaled to day 0
  geom_density(data = ind_surv, size = 1, alpha = 0.5, color = "steelblue4", fill = "steelblue4",
               aes(x = meta_date)) +
  facet_grid(sync_dfly ~ mean_dfly , scales = "free_y") + 
  xlab("date") + ylab("proportion") +
  theme(strip.text.x = element_text(size = 14),
        strip.text.y = element_text(size = 14)) 
plot6

```

# Growth rates by hatching date for all treatments

```{r, echo = F}

# even if they beat resource by 6 days
plot7 <- ggplot(ind, aes(x = hatch_date, y = growth_time, color = species)) + theme_bw() +
  geom_point(alpha = 0.65, size = 2) +
  stat_smooth(method = "lm", formula = y ~ poly(x, 2), se = T, size = 2) +
  facet_grid(sync_dfly ~ mean_dfly, scales = "free") + 
  theme(#legend.position = "none",
        strip.text.x = element_text(size = 14),
        strip.text.y = element_text(size = 14)) +
  xlab("hatch date") + ylab("growth rate") +
  scale_color_manual(values = c("#F28335", "#489C92"))
plot7

```

# Probability of metamorphosis by hatching date

```{r, echo = F}

plot9 <- ggplot(subset(ind, subset = (species == "dfly")), 
                aes(x = hatch_date, y = meta, color = sync_dfly)) +
  geom_point(alpha = 0.25) + theme_bw() +
  stat_smooth(method = 'glm', method.args = 'binomial', size = 2, alpha = 0.5, se = F) +
  facet_wrap(~mean_dfly) +
  scale_color_manual(values = wes_palette(n = 3, name = "GrandBudapest1")) 
plot9  
```

