---
title: "Phenology ABM: Analyses for Revisions"
author: "Shannon Carter"
date: "July 10, 2020"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: united
    highlight: tango 
editor_options: 
  chunk_output_type: console
---

```{r setup, include = F}
knitr::opts_chunk$set(echo = T, warning = F, message = F)

## Set working directory and clear memory
setwd("/Users/shannoncarter/Desktop/GitHub_Repos/phenology-abm")
rm(list = ls(all = T))

## Load required packages
library(tidyverse)
library(lme4)

## Load universal plotting elements
mytheme <-   theme_bw(base_size = 15, base_family = "Trebuchet MS") +
  theme(legend.text = element_text(size = 10),
        legend.title = element_text(size = 11),
        text = element_text(size = 14),
        axis.title = element_text(size = 12),
        axis.text  = element_text(size = 10, family = "Andale Mono"),
        panel.grid = element_blank())
```

# 2-Species Models 

## Load data
This has already been processed and cleaned in another script. This data has one row per individual and gives info on treatment parameters (11 levels of synchrony, 3 mean treatments), results of the simulation (number of survivors, biomass), and stats for the individual (hatching date, survival status, development time, final size).

```{r}
ind <- read.csv("ind2sp.csv", header = T)
head(ind)
```

## Survival model

#### Prepare data

Essentially, pool replicate treatments, filter for focal species, and calcuate standardized metrics for each treatment.

```{r}
## Prepare data
survival_model_data <- ind %>% 
  # remove run_num to pool all replicates of a treatment. also remove others we don't need
  select(-c(X, run_num, biom_fish, biom_dfly, biom_tot, end_date, growth_rate, id, sync_dfly, mean_dfly)) %>% 
  # "fish" is focal species that we manipulated
  filter(sp == "f") %>% 
  # only a few of these... I think the simulation ended before they "hatched"
  filter(!is.na(max_size)) %>% 
  # group by treatment to do the relativazation calculations
  group_by(sync_fish, mean_fish) %>% 
  # standardize arrival date for each treatment combination
  mutate(trt_max_hatch_date = max(hatch_date),
         trt_min_hatch_date = min(hatch_date),
         trt_mean_hatch_date = mean(hatch_date),
         hatch_percentile = round((hatch_date - trt_mean_hatch_date)/(trt_max_hatch_date - trt_min_hatch_date), 3))

head(survival_model_data)
```

#### Model survival

Logistic regression model with survival as a binary outcome

```{r}
## Fitness = survival model
m_surv <- glm(data = survival_model_data, meta ~ hatch_percentile + sync_fish + mean_fish +
                   hatch_percentile*sync_fish +
                   hatch_percentile*mean_fish +
                   sync_fish*mean_fish +
                   hatch_percentile*sync_fish*mean_fish,
              family = binomial(link = "logit"))

summary(m_surv)
coef(m_surv)

# make a data frame of individuals to add predicted survival to
predict_df <- data.frame(hatch_percentile = survival_model_data$hatch_percentile,
                sync_fish = survival_model_data$sync_fish,
                mean_fish = survival_model_data$mean_fish)

# calculate predicted survival + new value for each individual
m2 <- predict_df %>% 
  mutate(predicted_survival = predict(m_surv, type = "response")) %>% 
  mutate(coef = predicted_survival * (1 - predicted_survival))

# histogram of predicted survival
ggplot(m2, aes(x = predicted_survival)) + geom_histogram()
ggplot(m2, aes(x = coef)) + geom_histogram()

# calculate selection coefficients -- absolute and standardized
selection_coefficients <- m2 %>% 
  summarize(selection_coef = mean(coef), # absolute: 0.0628
            standardized_selection_coef = selection_coef/mean(survival_model_data$meta)) # standardized: 0.110

selection_coefficients
```


## Mass model

#### Prepare data

Almost the same as before, except we also filter out individuals that did not survive, and calculate relative size in each treatment

```{r}
## Prepare data
size_model_data <- ind %>% 
  # remove run_num to pool all replicates of a treatment. also remove others we don't need
  select(-c(X, run_num, biom_fish, biom_dfly, biom_tot, end_date, growth_rate, id, sync_dfly, mean_dfly)) %>% 
  # "fish" is focal species that we manipulated
  filter(sp == "f") %>% 
  # filter only those that survived
  filter(meta == 1) %>% 
  # only a few of these... I think the simulation ended before they "hatched"
  filter(!is.na(max_size)) %>% 
  # group by treatment to do the relativazation calculations
  group_by(sync_fish, mean_fish) %>% 
  # standardize fitness by dividing mass by treatment mean
  mutate(size_relative = max_size / mean(max_size)) %>% 
  # standardize arrival date for each treatment combination
  mutate(trt_max_hatch_date = max(hatch_date),
         trt_min_hatch_date = min(hatch_date),
         trt_mean_hatch_date = mean(hatch_date),
         hatch_percentile = round((hatch_date - trt_mean_hatch_date)/(trt_max_hatch_date - trt_min_hatch_date), 3))

head(size_model_data)

# View distribution of final sizes - pretty normal
ggplot(size_model_data, aes(x = size_relative)) + geom_histogram() + mytheme

```

#### Model mass

```{r}
## Fitness = mass model
m_size <- glm(data = size_model_data,
              size_relative ~
                hatch_percentile + sync_fish + mean_fish +
                hatch_percentile*sync_fish +
                hatch_percentile*mean_fish +
                sync_fish*mean_fish +
                hatch_percentile*sync_fish*mean_fish)
summary(m_size)
coef(m_size)
```

#### Plotting

```{r}
# adjust data for plotting-- remove some sync levels and put factors in order
m2_fewsync <- m2 %>% 
  filter(sync_fish == 3 | sync_fish == 15 | sync_fish == 27) %>% 
  mutate(synchrony_fish = case_when(sync_fish == 3 ~ "high",
                                    sync_fish == 15 ~ "medium",
                                    sync_fish == 27 ~ "low")) %>% 
  mutate(synchrony_fish = factor(synchrony_fish, levels = c("low", "medium", "high")), 
         mean_fish = factor(mean_fish, levels = c("early", "same", "late")))

ggplot(data = m2_fewsync, aes(x = hatch_percentile, y = predicted_survival, color = mean_fish, fill = mean_fish, shape = mean_fish)) +
  geom_point() + 
  geom_smooth() + #method = "lm", formula = y ~ poly(x, 3)) +
  facet_wrap(~synchrony_fish) +
  ylim(0, 1) +
  mytheme +
  labs(color = "relative mean arrival\nof focal species",
       fill = "relative mean arrival\nof focal species",
       shape = "relative mean arrival\nof focal species",
       x = "relative hatching date",
       y = "predicted survival") +
  scale_color_manual(values = c("#01A08A",  "#F4AC00", "#FF2401")) +
  scale_fill_manual(values = c("#01A08A",  "#F4AC00", "#FF2401")) +
  scale_shape_manual(values = c(21, 24, 23))
  
```

# 1-Species Models

## Load data
Very similar to the 2-species dataset. This data has one row per individual and gives info on treatment parameters (11 levels of synchrony, 3 levels of competitive asymmetry), results of the simulation (number of survivors, biomass), and stats for the individual (hatching date, survival status, development time, final size).

```{r}
ind_1 <- read.csv("ind1sp.csv", header = T)
head(ind_1)
```

## Survival model

#### Prepare data

Essentially, pool replicate treatments and calcuate standardized metrics for each treatment.

```{r}
## Prepare data
survival_model_data_1 <- ind_1 %>% 
  # remove run_num to pool all replicates of a treatment. also remove others we don't need
  select(-c(X, run_num, synchrony, asymmetry, n_surv, n_dead, biomass, mean_size, end_date, growth_time, growth_rate)) %>% 
  # group by treatment to do the relativazation calculations
  group_by(sync, asym) %>% 
  # standardize arrival date for each treatment combination
  mutate(trt_max_hatch_date = max(hatch_date),
         trt_min_hatch_date = min(hatch_date),
         trt_mean_hatch_date = mean(hatch_date),
         hatch_percentile = round((hatch_date - trt_mean_hatch_date)/(trt_max_hatch_date - trt_min_hatch_date), 3))

head(survival_model_data_1)
```

#### Model survival

Logistic regression model with survival as a binary outcome

```{r}
m_surv1 <- glm(data = survival_model_data_1, 
              meta ~ 
                hatch_percentile + sync + asym +
                hatch_percentile*sync +
                hatch_percentile*asym +
                sync*asym +
                hatch_percentile*sync*asym,
              family=binomial(link="logit"))
print(summary(m_surv1))
coef(m_surv1)

# make a data frame of individuals to add predicted survival to
predict_df1 <- data.frame(hatch_percentile = survival_model_data_1$hatch_percentile,
                sync = survival_model_data_1$sync,
                asym = survival_model_data_1$asym)

# calculate predicted survival + new value for each individual
m1 <- as_tibble(predict(m_surv1, predict_df1, type = "response")) %>% 
  mutate(coef = value * (1 - value))
m1 <- cbind(predict_df1, m1)

# calculate selection coefficients -- absolute and standardized
selection_coefficients_1sp <- m1 %>% 
  summarize(selection_coef = mean(coef, na.rm = T), # absolute: 0.06667
            standardized_selection_coef = selection_coef/mean(survival_model_data_1$meta, na.rm = T)) # standardized: 0.0889

selection_coefficients_1sp

```

## Mass model

#### Prepare data

Almost the same as before, except we also filter out individuals that did not survive, and calculate relative size in each treatment

```{r}
# Prepare data
size_model_data_1 <- ind_1 %>% 
  # remove run_num to pool all replicates of a treatment. also remove others we don't need
  select(-c(X, run_num, synchrony, asymmetry, n_surv, n_dead, biomass, mean_size, end_date, growth_time, growth_rate)) %>% 
  # group by treatment to do the relativazation calculations
  group_by(sync, asym) %>% 
  # standardize fitness by dividing mass by treatment mean
  mutate(size_relative = final_size / mean(final_size)) %>% 
  # standardize arrival date for each treatment combination
  mutate(trt_max_hatch_date = max(hatch_date),
         trt_min_hatch_date = min(hatch_date),
         trt_mean_hatch_date = mean(hatch_date),
         hatch_percentile = round((hatch_date - trt_mean_hatch_date)/(trt_max_hatch_date - trt_min_hatch_date), 3))

head(size_model_data_1)

# View distribution of final sizes - pretty normal
ggplot(size_model_data_1, aes(x = size_relative)) + geom_histogram() + mytheme

```

#### Model mass

```{r}
## Fitness = mass model
m_size_1 <- glm(data = size_model_data_1,
              size_relative ~
                hatch_percentile + sync + asym +
                hatch_percentile*sync +
                hatch_percentile*asym +
                sync*asym +
                hatch_percentile*sync*asym)
summary(m_size_1)
coef(m_size_1)
```

#### Plotting

```{r}
# adjust data for plotting-- remove some sync levels and put factors in order
m1_fewsync <- m1 %>% 
  filter(sync == 3 | sync == 15 | sync == 27) %>% 
  mutate(synchrony = case_when(sync == 3 ~ "high",
                                    sync == 15 ~ "medium",
                                    sync == 27 ~ "low")) %>% 
  mutate(asymmetry = case_when(asym == 0 ~ "none", 
                               asym == 0.5 ~ "weak",
                               asym == 1 ~ "strong")) %>% 
  mutate(synchrony = factor(synchrony, levels = c("low", "medium", "high")), 
         asymmetry = factor(asymmetry, levels = c("none", "weak", "strong")))

ggplot(data = m1_fewsync, aes(x = hatch_percentile, y = value, color = asymmetry, fill = asymmetry)) +
  geom_point() + 
  geom_smooth() + #method = "lm", formula = y ~ poly(x, 3)) +
  facet_wrap(~synchrony) +
  ylim(0, 1) +
  mytheme +
  labs(color = "competitive asymmetry",
       fill ="competitive asymmetry",
       x = "relative hatching date",
       y = "predicted survival") +
scale_color_manual(values = c("#acbab3",  "#fbd364", "#446353")) +
  scale_fill_manual(values = c("#acbab3",  "#fbd364", "#446353")) 
  
```