---
title: "Netlogo to R Pipeline: Resource Competition"
author: "Shannon Carter"
date: "March 11, 2019"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: united
    highlight: tango 
editor_options: 
  chunk_output_type: console
---

```{r setup, include = F}
knitr::opts_chunk$set(echo = T, warning = F, message = F)

## Set working directory and clear memory
setwd("/Users/shannoncarter/Desktop/GitHub_Repos/phenology-abm")
rm(list = ls(all = T))

## Load required packages
library(tidyverse)
library(lme4)

## Load universal plotting elements
mytheme <-   theme_bw(base_size = 15, base_family = "Trebuchet MS") +
  theme(legend.text = element_text(size = 10),
        legend.title = element_text(size = 11),
        text = element_text(size = 14),
        axis.title = element_text(size = 12),
        axis.text  = element_text(size = 10, family = "Andale Mono"),
        panel.grid = element_blank())
```

# Load Data
This has already been processed and cleaned in another script. This data has one row per individual and gives info on treatment parameters (11 levels of synchrony, 3 mean treatments), results of the simulation (number of survivors, biomass), and stats for the individual (hatching date, survival status, development time, final size).

```{r setup, echo = F, include = F}

# Processing to creat these dfs is done below, but can load these if working w same BS
ind <- read.csv("ind2sp.csv", header = T)
head(ind, 10)

```

# Survival Model

## Prepare data

Essentially, pool replicate treatments, filter for focal species, and calcuate standardized metrics for each treatment.

```{r}
## Prepare data
survival_model_data <- ind %>% 
  # remove run_num to pool all replicates of a treatment. also remove others we don't need
  select(-c(X, run_num, biom_fish, biom_dfly, biom_tot, end_date, growth_rate, id, sync_dfly, mean_dfly)) %>% 
  # "fish" is focal species that we manipulated
  filter(sp == "f") %>% 
  # only a few of these... I think the simulation ended before they "hatched"
  filter(!is.na(max_size)) %>% 
  # group by treatment to do the relativazation calculations
  group_by(sync_fish, mean_fish) %>% 
  # standardize arrival date for each treatment combination
  mutate(trt_max_hatch_date = max(hatch_date),
         trt_min_hatch_date = min(hatch_date),
         trt_mean_hatch_date = mean(hatch_date),
         hatch_percentile = round((hatch_date - trt_mean_hatch_date)/(trt_max_hatch_date - trt_min_hatch_date), 3))

head(survival_model_data, 10)
```

## Model

Logistic regression model with survival as a binary outcome

```{r}
## Fitness = survival model
m_surv <- glm(data = survival_model_data, meta ~ hatch_percentile + sync_fish + mean_fish +
                   hatch_percentile*sync_fish +
                   hatch_percentile*mean_fish +
                   sync_fish*mean_fish +
                   hatch_percentile*sync_fish*mean_fish,
              family=binomial(link="logit"))
print(summary(m_surv))
coef(m_surv)
```


# Mass Model

## Prepare data

Almost the same as before, except we also filter out individuals that did not survive, and calculate relative size in each treatment

```{r}
## Prepare data
size_model_data <- ind %>% 
  # remove run_num to pool all replicates of a treatment. also remove others we don't need
  select(-c(X, run_num, biom_fish, biom_dfly, biom_tot, end_date, growth_rate, id, sync_dfly, mean_dfly)) %>% 
  # "fish" is focal species that we manipulated
  filter(sp == "f") %>% 
  # filter only those that survived
  filter(meta == 1) %>% 
  # only a few of these... I think the simulation ended before they "hatched"
  filter(!is.na(max_size)) %>% 
  # group by treatment to do the relativazation calculations
  group_by(sync_fish, mean_fish) %>% 
  # standardize fitness by dividing mass by treatment mean
  mutate(size_relative = max_size / mean(max_size)) %>% 
  # standardize arrival date for each treatment combination
  mutate(trt_max_hatch_date = max(hatch_date),
         trt_min_hatch_date = min(hatch_date),
         trt_mean_hatch_date = mean(hatch_date),
         hatch_percentile = round((hatch_date - trt_mean_hatch_date)/(trt_max_hatch_date - trt_min_hatch_date), 3))

head(size_model_data, 10)

# View distribution of final sizes - pretty normal
ggplot(size_model_data, aes(x = size_relative)) + geom_histogram()

```

```{r}
## Fitness = mass model
m_size <- glm(data = size_model_data, 
              size_relative ~
                hatch_percentile + sync_fish + mean_fish +
                hatch_percentile*sync_fish +
                hatch_percentile*mean_fish +
                sync_fish*mean_fish +
                hatch_percentile*sync_fish*mean_fish)
print(summary(m_size))
coef(m_size)
```

